---
title: "Contributing to sAMRat"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Contributing to sAMRat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Welcome Contributors!

Thank you for your interest in contributing to **sAMRat** (Shiny Antimicrobial Resistance Analysis Tool)! This package is in active development, and we welcome contributions from the community.

> **📄 For the complete and most up-to-date contributing guidelines, please see our [CONTRIBUTING.md](https://github.com/Gero1999/sAMRat/blob/main/CONTRIBUTING.md) file in the repository.**

## Quick Overview

This page provides a quick overview of how to contribute. For detailed information, always refer to the official [CONTRIBUTING.md](https://github.com/Gero1999/sAMRat/blob/main/CONTRIBUTING.md) document.

## How to Contribute

### 1. Reporting Issues

Found a bug or have a feature request? 

- Visit our [GitHub Issues](https://github.com/Gero1999/sAMRat/issues) page
- Check if a similar issue already exists
- If not, create a new issue with:
  - Clear description of the problem or feature
  - Minimal reproducible example (for bugs)
  - Expected vs. actual behavior
  - System information (R version, OS, etc.)

### 2. Making Code Contributions

#### Getting Started

1. **Fork the repository** on GitHub
2. **Clone your fork** locally:
   ```bash
   git clone https://github.com/YOUR-USERNAME/sAMRat.git
   cd sAMRat
   ```
3. **Create a branch** with a descriptive name:
   ```bash
   git checkout -b X-feature/your-short-desc
   # or
   git checkout -b X-fix/your-short-desc
   ```
   Replace `X` with the respective issue number.

#### Development Workflow

1. **Make your changes** in small, focused commits
2. **Document your code** using roxygen2 comments
3. **Add tests** for new functionality
4. **Run checks** to ensure everything works:

```{r eval=FALSE}
# Load the package
devtools::load_all()

# Run tests
devtools::test()

# Run package checks
devtools::check()

# Update documentation
roxygen2::roxygenise()
```

#### Coding Style

- Follow **tidyverse style** for R code
- Keep functions small and focused
- Use meaningful variable and function names
- Comment complex logic
- Maintain consistent indentation

### 3. Submitting Pull Requests

When you're ready to submit your changes:

1. **Push your branch** to your fork
2. **Open a Pull Request** against the `main` branch
3. **Describe your changes**:
   - What problem does it solve?
   - How did you test it?
   - Any breaking changes?
4. **Wait for review** - maintainers will review and provide feedback

#### Pull Request Checklist

Before submitting, ensure:

- [ ] PR targets the `main` branch
- [ ] Code follows the project style
- [ ] Tests added/updated and passing
- [ ] Documentation updated
- [ ] `devtools::check()` passes with no new errors or warnings
- [ ] Commit messages are clear and descriptive

## Types of Contributions Needed

We welcome contributions in many areas:

### Code Contributions

- **New Features**: Additional analysis methods, visualizations, or export formats
- **Bug Fixes**: Fix reported issues or edge cases
- **Performance**: Optimize existing code for better performance
- **Refactoring**: Improve code structure and maintainability

### Documentation Contributions

- **Improve Vignettes**: Enhance existing tutorials or add new ones
- **Function Documentation**: Improve roxygen2 documentation
- **Examples**: Add practical examples to function documentation
- **Website**: Enhance the pkgdown website

### Testing Contributions

- **Unit Tests**: Increase test coverage
- **Integration Tests**: Test component interactions
- **Edge Cases**: Test boundary conditions and error handling

## Development Environment Setup

### Required Tools

```{r eval=FALSE}
# Install development tools if not already available
pkgs <- c("devtools", "testthat", "roxygen2", "pkgdown", "lintr")
install.packages(pkgs[!pkgs %in% installed.packages()[,"Package"]])
```

### Running the Application Locally

```{r eval=FALSE}
# Load the package
devtools::load_all()

# Run the Shiny app
shiny::runApp(system.file("shiny/app", package = "sAMRat"))
```

## Code Review Process

All contributions go through code review:

1. **Automated Checks**: CI/CD runs automated tests and checks
2. **Maintainer Review**: Core maintainers review the code
3. **Feedback Loop**: Address any requested changes
4. **Merge**: Once approved, your PR will be merged

## Community Guidelines

Please follow our [Code of Conduct](https://github.com/Gero1999/sAMRat/blob/main/CODE_OF_CONDUCT.md):

- Be respectful and inclusive
- Welcome newcomers
- Focus on constructive feedback
- Assume good intentions

## Getting Help

Need help contributing?

- **Questions**: Open a [GitHub Discussion](https://github.com/Gero1999/sAMRat/discussions) or issue
- **Email**: Contact the maintainer listed in [DESCRIPTION](https://github.com/Gero1999/sAMRat/blob/main/DESCRIPTION)
- **Documentation**: Read the [Architecture](architecture.html) guide to understand the codebase

## Recognition

Contributors are recognized in:

- The package DESCRIPTION file (for significant contributions)
- Release notes and changelogs
- The project README

## Resources

### Internal Resources

- [Architecture Guide](architecture.html) - Understand the codebase structure
- [Getting Started](getting-started.html) - Learn how to use the package
- [Official CONTRIBUTING.md](https://github.com/Gero1999/sAMRat/blob/main/CONTRIBUTING.md) - Complete contribution guidelines

### External Resources

- [R Packages Book](https://r-pkgs.org/) - Comprehensive guide to R package development
- [Tidyverse Style Guide](https://style.tidyverse.org/) - R coding style guidelines
- [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html) - Official R documentation
- [Advanced R](https://adv-r.hadley.nz/) - Deep dive into R programming

## Thank You!

Every contribution, no matter how small, helps improve sAMRat. We appreciate your time and effort in making this project better for everyone!

---

**Remember**: For the most current and detailed contribution guidelines, always refer to the [CONTRIBUTING.md](https://github.com/Gero1999/sAMRat/blob/main/CONTRIBUTING.md) file in the repository.
